---
title: ""
author: "Dasha Metropolitansky"
date: "April 1, 2019"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(readr)
library(stringr)
library(sf)
library(fs)
library(ggplot2)
library(gt)
library(lubridate)
library(ggthemes)
library(gganimate)
library(transformr)

#Reading in the RDS file with data from Hartford, CT

hartford <- read_rds(url("https://stacks.stanford.edu/file/druid:tr137st9964/tr137st9964_ct_hartford_2019_02_25.rds"))

#Downloading and unzipping the Shapes file with data from Hartford, CT

download.file(url = "https://stacks.stanford.edu/file/druid:tr137st9964/tr137st9964_ct_hartford_shapefiles_2019_02_25.tgz",
              destfile = "shapes.tgz") 
untar("shapes.tgz")
shapes_data = read_sf("ct_hartford_shapefiles/Hartford_Neighborhoods.shp")

#Deleting the files to clear up space and prevent duplicate downloads if running the code more than once

file_delete(c("shapes.tgz", "ct_hartford_shapefiles/"))
```

### Answering questions about the data
```{r include = FALSE}

#Arrest_made is a logical and since TRUE = 1 and FALSE = 0, this code totals the cases where an arrest was made. 

q1_1 = sum(hartford$arrest_made)
```
There were a total of `r q1_1` arrests made in Hartford.

```{r include = FALSE, message = FALSE}

#Filtering the Hartford data by cases where the arrestee is white, younger than 30, and a citation was issued

q1_2 = hartford %>%
  filter(subject_race == "white",
         subject_age < 30,
         citation_issued == "TRUE") 

#Calculating the length of the list of unique locations associated with the above subsetted dataframe

length(unique(q1_2$location))
```
There are `r length(unique(q1_2$location))` unique locations at which white individuals less than age 30 were issued citations.

```{r echo = FALSE}

#Filtering the Hartford data by cases where the arrestee is female and an arrest was made in the South End district

q1_3 = hartford %>% 
  filter(subject_sex == "female",
         district == "SOUTH END",
         arrest_made == "TRUE")

#Formatting the date in the style of "March 27, 2019"

q1_3$date = gsub(" 0", " ", format(q1_3$date, format = "%B %d, %Y"))

#Arranging by date so that the top result will be the first arrest

q1_3 = q1_3 %>%
  arrange(date)
```
The first arrest of a female in the South End district was on `r q1_3[1, 2]`.

```{r echo = FALSE}

#Filtering for districts that have "SOUTH" in their name

q1_4 = hartford %>% 
  filter(str_detect(string = district, pattern = "SOUTH")) %>%
  
#Grouping the data by officer, counting the number of stops for each group, and arranging in descending order so that the top result will have the greatest number of stops
  
  group_by(officer_id_hash) %>%
  count() %>% 
  arrange(desc(n))
```
Officer `r q1_4[1,1]` has made the most stops in districts with “South” in their name.

### Exploring the differences in arrest rates between races and genders

```{r echo = FALSE}
hartford %>%
  
#Grouping the data by subject sex, subject race, and whether an arrest was made, and counting the number of stops for each group
  
  group_by(subject_sex, subject_race, arrest_made) %>% 
  count() %>%
  
#Creating a "wider" dataset by creating two columns for the number of cases where arrest_made = FALSE and arrest_made = TRUE
  
  spread(key = arrest_made, value = n, fill = 0) %>%
  
#Creating an arrest_rate variable, which is the number of TRUEs/total number of stops for each race-gender grouping
  
  mutate(total = `FALSE` + `TRUE`, arrest_rate = `TRUE`/total) %>% 
  ungroup() %>%
  
#Renaming the levels of the factor 'subject_race'
  
  mutate(Race = recode_factor(subject_race,
                              "asian/pacific islander" = "Asian/Pacific Islander",
                              "black" = "Black",
                              "hispanic" = "Hispanic",
                              "other/unknown" = "Other/Unknown",
                              "white" = "White")) %>%
  
#Renaming the levels of the factor 'subject_sex'
  
  mutate(subject_sex = recode_factor(subject_sex, 
                                     "male" = "Male", 
                                     "female" = "Female")) %>%
  
#Selecting the relevant variables for the graph (arrest rate, race, and sex) and creating a table where the "subject_sex" column is replaced by "male" and "female" columns and arrest_rate is the value for each race/gender grouping rather than its own column 
  
  select(arrest_rate, Race, subject_sex) %>%
  spread(key = subject_sex, value = arrest_rate) %>%
  
#Formatting the graph: including percent signs next to the values, aligning the column text, and creating a title, subtitle, and source note
  
  gt() %>%
  fmt_percent(columns = vars("Male", "Female")) %>%
  tab_header(title = "Arrest Rate By Race and Gender in Hartford, CT from 2013-2016", 
             subtitle = "Percentage of traffic stops that result in arrests") %>%
  tab_spanner(label = "Gender", columns = c("Male", "Female")) %>%
  cols_align(columns = c("Male", "Female"), align = "center") %>%
  tab_source_note(md("Source: Stanford Open Policing Project")) 
```

### Visualizing the frequency of traffic stops over 24 hours

```{r echo = FALSE}
hartford %>% 
  
#Filtering out cases where the time of the stop is unknown
  
  filter(!is.na(time)) %>%
  
#Creating a histogram that shows how frequent stops are at different times
  
  ggplot(aes(x = time)) + 
  geom_histogram(bins = 300) + 
  
#Adding a title, x-axis and y-axis labels, and a source note
  
  labs(title = "Traffic Stops by Time of Day in Hartford CT", 
       x = "Time (24-Hour Clock)",
       y = "Number of Stops",
       caption = "Source: Stanford Open Policing Data") +
  
#Applying the "classic" theme
  
  theme_classic()
```

### Mapping the location of arrests by race and gender 

```{r echo = FALSE}
arrests_race = hartford %>% 
  
#Filtering for cases where an arrest was made
  
  filter(arrest_made == TRUE) %>% 
  
#Filtering out cases where the longitude, latitude, and district of where the arrest took place are unknown
  
  filter(!is.na(lng), !is.na(lat)) %>%
  filter(!is.na(district)) %>%
  
#Renaming the levels of the factor 'subject_race'
  
  mutate(subject_race = recode_factor(subject_race,
                              "asian/pacific islander" = "Asian/Pacific Islander",
                              "black" = "Black",
                              "hispanic" = "Hispanic",
                              "white" = "White")) %>%
  
#Renaming the levels of the factor 'subject_sex'
  
  mutate(subject_sex = recode_factor(subject_sex, 
                                     "male" = "Male", 
                                     "female" = "Female")) %>%
  
#Grouping by race
  
  group_by(subject_race)

#Converting the coordinates objects into a spatial object

arrest_locations = st_as_sf(arrests_race, 
                             coords = c("lng", "lat"), 
                             crs = 4326)

#Creating a map-shaped polygon using the location of each arrest as a point and colouring each point by the arrestee's gender

ggplot(data = shapes_data) +
  geom_sf() +
  geom_sf(data = arrest_locations, mapping = aes(fill = subject_sex, colour = subject_sex)) +
  
#Creating distinct maps for each race
  
  facet_grid(~subject_race) +
  theme_map() + 
  
#Including a title, subtitle, and source note
  
  labs(title = "Location of Arrests in Hartford, CT by Race and Gender",
       subtitle = "Arrests from 2013 - 2016", 
       caption = "Source: Stanford Open Policing Project" ) + 
  
#Moving the legend so that it doesn't cover the maps
  
  theme(legend.position = c(0.7, 1.1),
          legend.direction = "horizontal") +
  
#Modifying the legend colours
  
  scale_colour_manual(name = "Gender:", labels = c("Male", "Female"), values = c("royalblue2", "violetred1")) +
  scale_fill_manual(name = "Gender:", labels = c("Male", "Female"), values = c("royalblue2", "violetred1")) +
  
#Colouring the map gridlines white
  
  theme(panel.grid.major = element_line(colour = "white"))
```